# Coding Principles

This class is about two kinds of fundamental principles of coding. The first is fundamental methods of making code do what you want - if statement, loops, functions. The second is fundamental principles of good code. Although we are using R both kinds of principles hold when you write in other computer languages.

## Fundamental methods

1. [if statements](#if)
2. [loops](#loops)
3. [functions](#functions)



### If statements

So far we have written simple scripts that do things in order, top to bottom

```{r}
a <- 1 # define a variable
a <- a + 1 #add 1
echo(a) # output the result
```


### Loops{#loops}

Iterators


### Functions {#functions}

We need to add one *argument*, the p-value you want to report. The names you choose for the arguments are private to that argument, so it is not a problem if they conflict with other variables in your script. You put the arguments in the parentheses after `function` in the order you want them to default (just like the built-in functions you've used before). 

```{r}
report_p <- function(p) {
}
```

## Fundamental principles of good code

### Comment

Your most important collaborator is you from six months ago, and they don't answer email.


### Functionalise

Shorter code

A shortcut when writing code

A shortcut when reading code



### Avoid hard coded values

Transparency

Flexibility


### Ask for help

Writing a complete example

Your code is good enough to share