# Advanced Topics {-}

```{block, type="info"}
*Note this is a placeholder page. This material hasn't been taught yet. I am adding notes online as I can, so these pages in particular may evolve quickly*
```

The first ten topics of this course represent the core that I want every student to take away from the course. With these topics you should be equiped to focus on your personal data project for which makes up the majority of the grade on the course.

The remaining time in the timetable is available for supervision on this data project, and to cover more advanced topics. For these topics, rather than provide a lecture, we will work in small groups to follow exercises and tutorials which exist outside of the course. The motivation for this is two-fold. First, because all your future work will be teamwork, I want students to graduate from this course with experience working together on technical projects. Whether you are more or less confident in the technical requirements, you will learn a lot from trying to share what you know or think you know with a group. Second, a key skill for ongoing development in data science is to teach yourself. There world is rich in useful advice, tutorials and examples. Only by discovering how you can use these will you maximise your potential after you have finished this module.

## Interactive graphs with Shiny


### Before class

Please install shiny

```{r eval=FALSE}
install.packages("shiny")
install.packages("shinyjs")
install.packages("shinydashboard")
```
    
Please visit [shinyapps.io](https://www.shinyapps.io/) and make an account (you may be able to activate an account associated with your TUoS account by clicking 'sign in with Google')


### Exercises

We are going to follow [Lisa DeBruine](https://debruine.github.io/)'s tutorial -  [First Shiny App](https://psyteachr.github.io/shiny-tutorials/01-first-app.html).

Additional challenges / exercises  

* Make scatterplot rather than histogram
* Make the app load data which you save in the project directory
* Deploy to the web with [shinyapps.io](https://www.shinyapps.io/)
  * This page may be helpful when combining exercises 2 and 3: [Why does my app work locally, but not on shinyapps.io](https://support.rstudio.com/hc/en-us/articles/229848967-Why-does-my-app-work-locally-but-not-on-shinyapps-io-)
* Look for and share additional resources on shiny apps


### Resources

* This is part of a set of [Shiny Tutorials](https://psyteachr.github.io/shiny-tutorials/index.html)
* RStudio: [Shiny User Showcase](https://shiny.rstudio.com/gallery/)
* Example: [Movie-explorer](https://shiny.rstudio.com/gallery/movie-explorer.html)
* Example: [Enhanced sensitivity to group differences with decision modelling](https://sheffield-university.shinyapps.io/decision_power/)

### Examples of interactive graphs which use different tech (Javascript!)

 * [Interpreting Cohen's d Effect Size](https://rpsychologist.com/d3/cohend/) by Kristoffer Magnusson
 * Example: Grid histograms of [Constituencies by median age in 2017](https://olihawkins.com/visualisation/14) from Oli Hawkins

## Animated plots with gganimate

This is a package developed by Thomas Lin Pedersen and colleagues. It has it's own page here:
[gganimate](https://gganimate.com/

This is the documentation. It explains how to install it, and what you can do with it, and - if you dig a bit - what the options are for the various functions.

Get started by installing:


```{r, eval=FALSE}
install.packages("gganimate")
```

This may give you a warning about additional packages you can install to help create output files in gif or movie file format

```{r, eval=FALSE}
install.packages("gifski")
install.packages("av")
```

This *in turn* may ask for additional software to be installed on your machine, and the instructions will depende on what kind of operating system you are on.

Assuming you have fixed that, it's time to make some plots:

Here are three different turorials on gganimate:

* gganimate.com: [Getting Started](https://gganimate.com/articles/gganimate.html)
* datanovia.com: [How to Create Plots with Beautiful Animation in R](https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r)
* rpubs.com/jedoenriquez: [An Intro to Animating Charts with gganimate](https://rpubs.com/jedoenriquez/animatingchartsintro)

The first is written by the package authors. Pay attention the material on 'rendering' at the end, this is important if you want to make animations you can share.

The second gives a good exploration of the different option, nicely showing off possible effects and how you would use them with different data sets.

The third is the shortest and simplest, but note that they use a different structure for creating ggplot objects than we do (piping the data to ggplot). If you are not sure, start here.

There is some nuance in getting the animation to save. You are having trouble with this, I recommend you divide the data frame preparation, animate plot creation and rendering into seperate stages. Once you have made a plot object, `p`, try this

```{r, eval=FALSE}
animate(p,renderer=gifski_renderer("my_animation.gif"))
```

### Exercises

* Pick some data to animate
  * Control the speed of the animation
  * Save the animation as a gif file and add it to a RMarkdown document (and publish to github pages)
    * bonus question, why do it this way rather than use R to generate the animation on the page?
  * Use the `{frame_time}` attribute in a label
    * How does gganimate know what values to put in `frame_time`?
  * Add small black marks to show the starting values of the data. Which option does this
    * Look at the documentation and see if you can figure out what line to add to achieve this
      * [Hint](https://gganimate.com/reference/shadow_mark.html)
  * Why do you need to render a gif? 
      * Why can't you just right-click 'copy image' on the RStudio output? What happens if you do this?

## Database queries: SQL

  * We will look at this interactive tutorial https://sqlzoo.net/wiki/SELECT_basics



## Reproducibility

Dependencies

 * Package management 
    * https://groundhogr.com/
    * renv: https://github.com/rstudio/renv
    * e.g. https://luisdva.github.io/rstats/annotater/
 * Reproducible workflows
    * [Targets](https://books.ropensci.org/targets/) (was Drake)
 * Containers
    * e.g. [docker](https://www.docker.com/)



